Something about Performance, 2D Component of NEMO. (TBC)

### 2015-02-12 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
Add data and plot of best psykal (serial) results
Add performance profile table

Since any real oceanographic computational model must output results,
we ensure that any \psykal version of shallow retains the Input/Output
capability of the original. This aids in limiting the optimisations
that can be performed on the \psykal version to those that should also
be applicable to full oceanographic models. Note that although we
retain the I/O functionality, all of the results presented in this work
carefully exclude the effects of I/O since it is compute performance
that interests us here.

In order to maximise the flexibility (and thus potential for
optimisation) of the \psykal version of nemolite2d, we made the kernels
as fine-grained as possible. In this case, this resulted in eight
distinct kernels, each of which operated on a single field at a single
point (since we have chosen to use point-wise kernels). With a little
bit of tidying/re-structuring, we found it was possible to express the
contents of the main time-stepping loop as a single invoke (a call to
the PSy layer) and a call to the I/O system
(Figure~\ref{FIG_psykal_shallow_structure}). The single PSy-layer
routine then consists of applying each of the kernels to all of the
points on the model mesh requiring an update. In its basic,
unoptimised (`vanilla') form, this PSy-layer routine then contains a
doubly-nested loop around each kernel call, as indicated by the
pseudo-code in Figure~\ref{FIG_psykal_shallow_structure}.

\begin{figure}
\centering
%\includegraphics[width=85mm]{psykal_shallow}
\caption{A schematic of the vanilla \psykal version of the shallow code.}
\label{FIG_psykal_shallow_structure}
\end{figure}

@@

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
\caption{The matrix of compilers and CPUs used in this work. The use
  of a compiler on a given CPU is indicated by the specification of
  the version of the compiler in the relevent element. No entry
  indicates that a compiler was not available/used on that CPU.}
\label{TABLE_compilers}
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
                 & \multicolumn{4}{c|}{Compiler}             \\
\hline
                 & Gnu   & Intel       & Cray    & IBM bgxlf\\
\hline
Intel Haswell    & 4.8.2 & 14.0.0      &         &          \\
Intel Ivy Bridge & 4.9.1 & 14.0.1.106  & 8.3.3   &          \\
IBM A2 (BGQ)     &       &             &         & 14.1     \\
IBM Power 8      &       &             &         &          \\
\hline
\end{tabular}
\end{table}

@@

@@

\item In-lined field-copy operation

\item Manually in-line kernel bodies into middle-layer code

\item Plus others\ldots

@@

In Figure~\ref{FIG_orig_perf_summary} we plot the performance of the
original version of the nemolite2d code for the range of compilers and
hardware considered here.

\begin{figure}[!t]
\centering
%\includegraphics[width=2.8in]{orig_summary}
\caption{Summary of the performance of the original version of the 
nemolite2d code on 
the range of compilers and hardware under consideration.}
\label{FIG_orig_perf_summary}
\end{figure}

Moving now the to the \psykal version of nemolite2d,
Figure~\ref{FIG_best_psykal_perf_summary} plots the performance of the
fastest \psykal version for each of the compiler/CPU combinations. The
only significant difference between this and the plot of the
performance of the original version of the code in
Figure~\ref{FIG_orig_perf_summary} is that the difference between the
Gnu- and Intel-compiled executables is slightly more marked. This is
clarified in Figure~\ref{FIG_slowdown_summary} which plots the
percentage difference between the performance of the original and the
(best) \psykal versions of shallow for each compiler/CPU combination.

\begin{figure}[!t]
\centering
%\includegraphics[width=7.5cm]{best_psykal_summary}
\caption{Summary of the best performance achieved by any \psykal 
version of shallow with each of the compilers and CPUs under 
consideration.}
\label{FIG_best_psykal_perf_summary}
\end{figure}

\begin{figure}[!t]
\centering
%\includegraphics[width=2.8in]{slowdown_summary}
\caption{Comparison of the performance of the best \psykal
version with that of the original version of the code. A negative value 
indicates that the \psykal version is faster than the original.}
\label{FIG_slowdown_summary}
\end{figure}

Having show that, in general, we can recover and occasionally improve
upon the performance of the original version of shallow, the next
logical step is to examine the necessary code transformations.  We do
this for the $256^{2}$ case since this fits within cache on both the
Haswell and Ivy Bridge CPUs we are using here.
Table~\ref{TABLE_opt_breakdown} shows detailed timings for this case
after each transformation has been applied to the code. The same data
is visualised in Figure~\ref{FIG_opt_stages_256} where the performance
of the \psykal version for a given compiler/CPU is given relative to
the performance of the original with the same compiler/CPU.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=95mm]{opt_stages_256}
\caption{The performance of the \psykal version of nemolite2d for the
  $256^{2}$ domain at each stage of optimisation. Results are given as
  a percentage of the performance of the original code for each
  compiler/CPU combination. A figure greater than 100\% indicates that
  the \psykal version performs better than the original.}
\label{FIG_opt_stages_256}
\end{figure}

>>>>

Since any real oceanographic computational model must output results,
we ensure that any \psykal version of nemolite2d retains the Input/Output
capability of the original. This aids in limiting the optimisations
that can be performed on the \psykal version to those that should also
be applicable to full oceanographic models. Note that although we
retain the I/O functionality, all of the results presented in this work
carefully exclude the effects of I/O since it is compute performance
that interests us here.

In order to maximise the flexibility (and thus potential for
optimisation) of the \psykal version of nemolite2d, we made the kernels
as fine-grained as possible. In this case, this resulted in eight
distinct kernels, each of which operated on a single field at a single
point (since we have chosen to use point-wise kernels). With a little
bit of tidying/re-structuring, we found it was possible to express the
contents of the main time-stepping loop as a single invoke (a call to
the PSy layer) and a call to the I/O system
(Figure~\ref{FIG_psykal_nemolite2d_structure}). The single PSy-layer
routine then consists of applying each of the kernels to all of the
points on the model mesh requiring an update. In its basic,
unoptimised (`vanilla') form, this PSy-layer routine then contains a
doubly-nested loop around each kernel call, as indicated by the
pseudo-code in Figure~\ref{FIG_psykal_nemolite2d_structure}.

\begin{figure}
\centering
%\includegraphics[width=85mm]{psykal_nemolite2d}
\caption{A schematic of the vanilla \psykal version of the nemolite2d code.}
\label{FIG_psykal_nemolite2d_structure}
\end{figure}

@@

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
\caption{The matrix of compilers and CPUs used in this work. The use
  of a compiler on a given CPU is indicated by the specification of
  the version of the compiler in the relevent element. No entry
  indicates that a compiler was not available/used on that CPU.}
\label{TABLE_compilers}
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
                 & \multicolumn{4}{c|}{Compiler}             \\
\hline
                 & Gnu   & Intel       & Cray    & IBM bgxlf\\
\hline
Intel Haswell    & 4.8.2 & 14.0.0      &         &          \\
Intel Ivy Bridge & 4.9.1 & 15.0.0.090  & 8.3.3   &          \\
IBM A2 (BGQ)     &       &             &         &   --     \\
IBM Power 8      &       &             &         &          \\
\hline
\end{tabular}
\end{table}

@@

\begin{table}[!t]
\caption{The performance profile of the original and \psykal 
versions of nemolite2d ($128^{2}$ domain and Intel compiler).}
\label{TABLE_profile}
\centering
\begin{tabular}{l|c|c|c|c}
\hline
Section & \multicolumn{2}{c|}{Original} & \multicolumn{2}{c}{\psykal} \\
        & Time (s) & \%  & Time (s) & \% \\
\hline
Continuity  & 0.096 & 2.7  & 0.33  & 6.6 \\
Momentum    & 3.00  & 84.5 & 4.05  & 79.5 \\
BCs         & 0.16  & 4.6  & 0.29  & 5.7 \\
Time-update & 0.29  & 8.1  & 0.41  & 8.1 \\
\hline
\end{tabular}
\end{table}

@@

\item Addition of 'safe-loop' directives

\item In-line Momentum kernel bodies into middle-layer code

\item Force SIMD vectorisation of the Momentum kernels using directives

\item Work around limitations related to derived types

\item In-line the Continuity kernel

\item In-line remaining kernels (BCs and time-update)

\item In-line field-copy operations

@@

In Figure~\ref{FIG_orig_perf_summary} we plot the serial performance
of the original version of the nemolite2d code for the range of
compilers and hardware considered here.

\begin{figure}[!t]
\centering
\includegraphics[width=2.8in]{orig_summary}
\caption{Summary of the performance of the original version of the 
nemolite2d code on 
the range of compilers and hardware under consideration.}
\label{FIG_orig_perf_summary}
\end{figure}

Moving now the to the \psykal version of nemolite2d,
Figure~\ref{FIG_best_psykal_perf_summary} plots the performance of the
fastest \psykal version for each of the compiler/CPU combinations. The
only significant difference between this and the plot of the
performance of the original version of the code in
Figure~\ref{FIG_orig_perf_summary} is that the difference between the
Gnu- and Intel-compiled executables is slightly more marked. This is
clarified in Figure~\ref{FIG_slowdown_summary} which plots the
percentage difference between the performance of the original and the
(best) \psykal versions of nemolite2d for each compiler/CPU combination.

\begin{figure}[!t]
\centering
\includegraphics[width=7.5cm]{best_psykal_summary}
\caption{Summary of the best performance achieved by any \psykal 
version of nemolite2d with each of the compilers and CPUs under 
consideration.}
\label{FIG_best_psykal_perf_summary}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=2.8in]{slowdown_summary}
\caption{Comparison of the performance of the best \psykal
version with that of the original version of the code. A negative value 
indicates that the \psykal version is faster than the original.}
\label{FIG_slowdown_summary}
\end{figure}

Having show that, in general, we can recover and occasionally improve
upon the performance of the original version of nemolite2d, the next
logical step is to examine the necessary code transformations.  We do
this for the $256^{2}$ case since this fits within cache on both the
Haswell and Ivy Bridge CPUs we are using here.
Table~\ref{TABLE_opt_breakdown} shows detailed timings for this case
after each transformation has been applied to the code. The same data
is visualised in Figure~\ref{FIG_opt_stages_256} where the performance
of the \psykal version for a given compiler/CPU is given relative to
the performance of the original with the same compiler/CPU.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=90mm]{opt_stages_256}
\caption{The performance of the \psykal version of nemolite2d for the
  $256^{2}$ domain at each stage of optimisation. Results are given as
  a percentage of the performance of the original code for each
  compiler/CPU combination. A figure greater than 100\% indicates that
  the \psykal version performs better than the original.}
\label{FIG_opt_stages_256}
\end{figure}
